---
title: "Lab 4a"
author: "Jas Sur"
date: "2/27/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1

Setup data for pressure, temperature and yield
```{r}
temp.data <- rep(c(150,160,170), each=6)
pressure.data <- rep(c(200,215,230), each=3, times=2)
yield <- c(90.4, 90.2, 90.7,90.6,90.2,90.4,90.1,90.3,90.5,90.6,89.9,90.1,90.5,90.7,90.8,90.9,90.4,90.1)
chem.process <- data.frame(temp.data, pressure.data, yield)
attach(chem.process)

chem.process$temp <-factor(temp.data, levels =c(150,160,170), labels=c("150C", "160C", "170C"))
chem.process$pressure <-factor(pressure.data, levels =c(200,215,230), labels=c("200Psig","215Psig", "230Psig"))

head(chem.process)
str(chem.process)
attach(chem.process)
```

### a. Analyze the data and draw conclusions, Use alpha = 0.05

a1. Plotting: boxplot of yield vs temp, yield vs pressure, yeld vs temp and pressure, interaction plot
```{r}

tapply(yield, list(temp,pressure), mean)
boxplot(yield~temp, main="yield vs temp")
boxplot(yield~pressure, main="yield vs pressure")
boxplot(yield~temp*pressure, main="yield vs temp*pressre")

interaction.plot(temp, pressure, yield)
interaction.plot(pressure, temp, yield)

```
Slopes are changing, so there is some kind of interaction


a2. Build a linear model using aov(). Print ANOVA output. Are main factors and interaction factor significant?

```{r}
chem.process.mod <- aov(yield~temp*pressure)
summary.aov(chem.process.mod)

summary.lm(chem.process.mod)

chem.process.nointeraction.mod <- aov(yield~temp+pressure)
anova(chem.process.mod, chem.process.nointeraction.mod)
```

Since P-value >.05, we accept null and reject Alternate. The interaction term is in-significant, so we can discard the interaction term. We take the partial model instead of full model


a3. Print out regression model with coefficients
```{r}

summary.lm(chem.process.nointeraction.mod)
```

Only the intercept is significant, so it means that the linear model can remove other terms while predicting the values. 


### b. Prepare appropriate residuals plot and comment on the model's adequecy
```{r}

summary.lm(chem.process.nointeraction.mod)
summary.aov(chem.process.nointeraction.mod)
```
It looks like the difference of mean for both temperature and pressure is the same, so there is not much difference becaue of temperatuer or pressure.

### c. Under what condition would you operate this process
```{r}
tukeys <- TukeyHSD(chem.process.nointeraction.mod)
tukeys
plot(tukeys)
```
Pvalue is large for both pressure, and temperature at all conditions, so we accept null hypothesis that there is not much difference due to the changing conditions

```{r}
library(agricolae)
lsd <- LSD.test(chem.process.nointeraction.mod, c("temp", "pressure"), console=T)
lsd
plot(lsd)
detach()
```

All terms generate almost the same yield, so we should choose the one which can be operatored at least cost. I would choose the one at 150C and 200Psig


## Question 2

setup data
```{r}
operator<- rep(c(1,2,3), each=8)
machine<- rep(c(1,2,3,4), each=2, times=3)

strength <- c(109, 110,110,115,108,109,110,108,110,112,110,111,111,109,114,112,116,114,112,115,114,119,120,117)

fiber <- data.frame(operator, machine, strength)

fiber$operator <-factor(operator, levels =c(1,2,3), labels=c("Op1", "Op2", "Op3"))
fiber$machine <-factor(machine, levels =c(1,2,3,4), labels=c("Mach1","Mach2","Mach3","Mach4"))
operator <-fiber$operator
machine <-fiber$machine

head(fiber)
str(fiber)
attach(fiber)

```

a1. Plotting: boxplot of strength vs operator, strength vs machine, strength vs operator and machine, interaction plot
```{r}

tapply(strength, list(operator,machine), mean)
boxplot(strength~operator, main="strength vs operator")
boxplot(strength~machine, main="strength vs machine")
boxplot(strength~operator*machine, main="strength vs operator*machine")

interaction.plot(operator, machine, strength)
interaction.plot(machine, operator, strength)

```
Interaction plots are intersecting, so there is a indication of strong interaction


a2. Build a linear model using aov(). Print ANOVA output. Are main factors and interaction factor significant?

```{r}
fiber.mod <- aov(strength~operator*machine)
summary.aov(fiber.mod)

summary.lm(fiber.mod)

operator
fiber.nointeraction.mod <- aov(strength~operator+machine)
anova(fiber.mod, fiber.nointeraction.mod)
```

Since P-value >.05, we accept null and reject Alternate. The interaction term is in-significant, so we can discard the interaction term. We take the partial model instead of full model


a3. Print out regression model with coefficients
```{r}

summary.lm(fiber.nointeraction.mod)
```
Intercept and Operator are both siginficant


### b. Prepare appropriate residuals plot and comment on the model's adequecy
```{r}

summary.lm(fiber.nointeraction.mod)
summary.aov(fiber.nointeraction.mod)
```
It looks like the difference of mean is same for machine while it's different for operator. We need to run Tukey's test to see which one are different 

### c. Under what condition would you operate this process
```{r}
tukeys <- TukeyHSD(fiber.nointeraction.mod)
tukeys
plot(tukeys)
```
Pvalue is large for both pressure, and temperature at all conditions, so we accept null hypothesis that there is not much difference due to the changing conditions

```{r}
library(agricolae)
lsd <- LSD.test(fiber.nointeraction.mod, c("operator", "machine"), console=T)
lsd
plot(lsd)
```
Operator3 and machine 4 appears to give the best strength

## Question 3. 
### a. create data frame with days and solution and growth as response variable

```{r}
solution <-rep(c(1,2,3), each=4)
days <-  rep(c(1,2,3,4), each=3)
solution <- factor(solution, levels=c(1,2,3), labels = c("Sol1", "Sol2", "Sol3"))
days <- factor(days, levels =c(1,2,3,4), labels = c("Day1", "Day2", "Day3", "Day4"))

growth<-c(13,22,18,39,16,24,17,44,5,4,1,22)

powder <-data.frame(solution, days, growth)

head(powder)
str(powder)
attach(powder)
```

Any evidence that the solution affect bacteria growth

```{r}
length(growth)
length(solution)
boxplot(growth~solution, main="growth vs solution")

interaction.plot(solution, days, growth)
interaction.plot(days, solution, growth)

```
The interaction appears to be weak cause these interaction plots are intersecting.


```{r}
powder.mod <- aov(growth~solution*days)
summary.aov(powder.mod)

summary.lm(powder.mod)

powder.nointeraction.mod <- aov(growth~solution+days)
anova(powder.mod, powder.nointeraction.mod)
```
Pvalue is 0, so there appears to be no interaction, so we chose non-interactive model.

### b. Perform a TukeyHSD to compare solution treatment means
```{r}
tukeys <- TukeyHSD(powder.nointeraction.mod)
plot(tukeys)
```
Difference of mean is the same so the solution or days doesn't affect much of the bacterial growth

```{r}
lsd <- LSD.test(powder.nointeraction.mod, c("solution", "days"), console = T)
plot(lsd)
```

The solution and days doesn't appear to have much affect on growth.

### c. Check assumptions of the residuals

```{r}

res <- residuals(powder.nointeraction.mod)

qqnorm(res)
qqline(res)
shapiro.test(res)
#check variance
bartlett.test(res~solution)
bartlett.test(res~days)
```

The data is normal and variance is equal.