---
title: "Lab0"
author: "Jas Sur"
date: "1/26/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```
# Lab 0 Problems

## Question 1.a
### Generate a random sample of 100 from t-distribtion, degree of freedom 10. 
### Check qqnorm(); qqline(), Shapiro test. Remarks
```{r q1.b}

sample.t <- rt(100, df=10)
qqnorm(sample.t)
qqline(sample.t)
shapiro.test(sample.t)

```
***
1. Null hypothesis (H0 sample data is normal)
2. Alternative hypothesis (H1 sample data is not normal)

Since the pvalue > 0.05, which means that the null hypothesis that values are normal is accepted and the alternative hypothesis is rejected.
***

## Question 1.b
### Generate a random sample of 100 from a Chi-square distribution, df =5. 
### Perform same procedures as in (a) . Remarks.

```{r q.1b}

sample.1b <-seq(-4, 4, 0.01)
sample.chi = dchisq(sample.1b, df=5)
qqnorm(sample.chi)
qqline(sample.chi)
shapiro.test(sample.chi)

```

***
1. Null hypothesis (H0 sample data is normal)

2. Alternative hypothesis (H1 sample data is not normal)

Since the pvalue < 0.05, which means that the null hypothesis that values are normal is rejected and the alternative hypothesis is accepted.


## Question 2.a
### Write a 95%-CI for the population mean. 

***
Confidence interval for population mean can be determined using
the following 
```{r q.2a}
sample <-c(26.4,23.5,25.4,22.9,25.2,39.2,25.5,31.9,26.0,44.6,35.5,38.6,
           30.1,31.0,30.8,32.8,47.7,39.1,55.3,50.7,73.8,71.1,68.4,77.1,
           19.4,19.3,18.7,19.0,23.2,21.3,23.2,19.9,18.9,19.8,19.6,21.9)
hist(sample)


n<-length(sample)
xbar<-mean(sample)
s<-sd(sample)
sprintf("mean %f, std dev %f, sample size %d", xbar, s, n)
margin <- qt(0.95,df=n-1)*s/sqrt(n)
low_ci <- xbar - margin

high_ci <- xbar + margin
sprintf("confidence interval for the population mean is (%f, %f)", low_ci, high_ci)
boxplot(sample, col = 3)
lines(c(0.75,1.25),c(low_ci,low_ci),col=4)
lines(c(0.75,1.25),c(xbar,xbar),col=2,lwd=2)
lines(c(0.75,1.25),c(high_ci,high_ci),col=4)
legend("topleft", c("95% CI", "Mean"), lty=1,col = c(4, 2),bty ="n")
```

***
95 % Confidence interval is reasonable compare to the sample mean

### What assumption about population for the work, suppose the sample is random.
***
The assumptions would be that the data is normal and the errors are constant. 


## Question 2.b
### Write a 95%- CI for population standard deviation.

***
Confidence interval for population standard deviation can be determined using
the following formula
confidence interval: $CI = [√(n-1)s^{2}/X^{2}(α/2), √(n-1)s^{2}/ X^{2}(1-α/2)]$

```{r q.2b}

sample <-c(26.4,23.5,25.4,22.9,25.2,39.2,25.5,31.9,26.0,44.6,35.5,38.6,
           30.1,31.0,30.8,32.8,47.7,39.1,55.3,50.7,73.8,71.1,68.4,77.1,
           19.4,19.3,18.7,19.0,23.2,21.3,23.2,19.9,18.9,19.8,19.6,21.9)
n <- length(sample)
xbar<-mean(sample)
s<-sd(sample)
sprintf("population sd %f, size %d, mean %f", s, n, xbar)

left <- qchisq(p=.05, df=n-1, lower.tail=FALSE)
right <- qchisq(p=.95, df=n-1, lower.tail=FALSE)
low_ci <- sqrt((n-1)*s^2/left)
high_ci <- sqrt((n-1)*s^2/right)

sprintf("Confidence interval for population std dev is  %f, %f", low_ci, high_ci)
boxplot(sample-s, col = 3)
lines(c(0.75,1.25),c(low_ci,low_ci),col=4)
lines(c(0.75,1.25),c(s,s),col=2,lwd=2)
lines(c(0.75,1.25),c(high_ci,high_ci),col=4)
legend("topleft", c("95% CI", "Std dev"), lty=1,col = c(4, 2),bty ="n")
```
***
A confidence interval for a standard deviation is a range of values that is likely to contain true population standard deviation with a certain level of confidence.
***

# Question 3
### Run the following code

```{r q.3}

sample <-c(26.4,23.5,25.4,22.9,25.2,39.2,25.5,31.9,26.0,44.6,35.5,38.6,
           30.1,31.0,30.8,32.8,47.7,39.1,55.3,50.7,73.8,71.1,68.4,77.1,
           19.4,19.3,18.7,19.0,23.2,21.3,23.2,19.9,18.9,19.8,19.6,21.9)
hist(sample)
sample.s <-sort(sample) #sort data increasing
rank <- rank(sample.s) #rank data from 1 to 36
size <- length(sample.s) # size of data
p <- (rank-.5)/size #cummulative prob of data
z.quantile <- qnorm(p) # Standard Normal quantiles with such probability

data <- cbind(rank, sample.s, p, z.quantile)
data

#scatterplot of x=Z quantiles, y= data sorted
plot(x=z.quantile, y=sample.s, pch=16, main="QQ Plot") 
abline(lm(sample.s ~ z.quantile))
```

The QQ plot suggests that the data is not normal
