---
title: "Lab 4b"
author: "Jas Sur"
date: "3/1/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
### a. Set up the data frame, named “chem”, with “Bolt” and “Chemical” factors, 
### “strength” as response.

```{r}

setup_q_1 <- function()
{ 
chemical.data <- rep(c(1,2,3,4), each=5)
bolt.data <- rep(c(1,2,3,4,5), times=4)

strength <- c(73,68,74,71,67,73,67,75,72,70,75,68,78,73,68,73,71,75,75,69)

effect <- data.frame(strength, chemical.data, bolt.data)
effect$chemical <- factor(chemical.data, levels=c(1,2,3,4), 
                          labels = c("Chemical1", "Chemical2", "Chemical3", 
                                     "Chemical4"))
effect$bolt <- factor(bolt.data, levels=c(1,2,3,4,5), 
                      labels=c("Bolt1", "Bolt2", "Bolt3", "Bolt4", "Bolt5"))
return (effect)
}
effect <-setup_q_1()
str(effect)
head(effect)
attach(effect)
```

### b.i. Any evidence that the Chemical affect Strength ? 
### Note: Chemical is the treatment factor. 

```{r}
tapply(strength, list(chemical, bolt), mean)
boxplot(strength~chemical, main="strength vs chemical")
boxplot(strength~bolt, main="strength vs bolt")
boxplot(strength~chemical*bolt, main="strength vs bolt*chemical")

interaction.plot(bolt, chemical, strength)
interaction.plot(chemical, bolt, strength)

```

> There appears to be strong interaction between bolt and chemical on it appears 
to be causing a varying response

```{r}
effect.interaction.mod <- aov(strength~chemical*bolt, data=effect)
summary.aov(effect.interaction.mod)
```

Since n=1 (single replicate) ; MSE =SSE/ab(n-1), MSE can’t be calculated!
We can assume no significant between chemical and bolt

We need to run Tukey1df to check if the interaction term is significant
```{r}
library(daewr)
tukey.interaction <- data.frame(strength, chemical, bolt)
Tukey1df(tukey.interaction)
```

> NonAdditivity is 0.7508, so interaction term is not significant


```{r}
effect.mod <- aov(strength~chemical+bolt, data=effect)
summary.aov(effect.mod)
```
> The Chemical has pvalue >0.05, so the null hypothesis that the chemical 
doesn't affect strength is accepted and the alternative hypothesis that the 
chemical affect strength is rejected

>The bolt has pvalue <0.05, so the null hypothesis that the chemical doesn't 
affect strength is rejected and the alternative hypothesis that the chemical 
affect strength is accepted

### b.ii. Is Bolt is an effective blocking factor?
> Since it's pvalue is significant, the blocking factor could be effective. 
We need to check if the MSE  increases or decreases without blocking factor. 
If it increases, then we know that Blocking factor is effective.

```{r}
effect.1.mod <- aov(strength~chemical, data=effect)
summary.aov(effect.1.mod)
```
> MSE increases significantly without bolt, so yes the bolt act as an effective 
blocking factor

### c. Perform a TukeyHSD to compare the treatment means. Which Chemical is the 
### preferred(bring the highest strength)

```{r}
tukey <- TukeyHSD(effect.mod)
tukey
plot(tukey)
```

Chemical difference of mean is not significant, so it's clear chemical doesn't 
affect the strength. Some of the bolt values have pvalue <0.05, so it's clearly 
affect the strength.

### d. Check the assumptions of the residuals.

```{r}
res <- residuals(effect.mod)

qqnorm(res)
qqline(res)
shapiro.test(res)
bartlett.test(res~bolt)
bartlett.test(res~chemical)
```

The data doesn't appear to be normal, but it has const variance.

```{r echo=FALSE}
detach()
```

***

## Question 2
### a. set up the data frame named eye, subject, distance as factors and time as response

```{r}
setup_q_2 <-function()
{
subject.data <-rep(c(1,2,3,4,5), time=4)
distance.data <- rep(c(4,6,8,10), each=5)
time <-c(10,6,6,6,6,7,6,6,1,6,5,3,3,2,5,6,4,4,2,3)
eye <- data.frame(time, subject.data, distance.data)

eye$subject <- factor(subject.data, c(1,2,3,4,5), label=c("Sub1","Sub2", "Sub3","Sub4", "Sub5"))

eye$distance <- factor(distance.data, c(4,6,8,10), label=c("Dist4","Dist6", "Dist8","Dist10"))
return (eye)
}
eye <- setup_q_2()
head(eye)
str(eye)
```
### b. build a regression model named eye.mod. Is "subject" an effecting blocking factor?
```{r}
attach(eye)
eye.mod <- aov(time~subject*distance, data=effect)
summary.aov(eye.mod)
```
Since n=1, so there are no pvalues. We need to check it using Tukeys1df

```{r}
library(daewr)
tukey.interaction <- data.frame(time, subject, distance)
Tukey1df(tukey.interaction)
```

NonAdditivity is 0.6624, so interaction is not significant.


```{r}
eye.mod <- aov(time~subject+distance, data=effect)
summary.aov(eye.mod)
```
It's evident that both subject and distance are both significant.


```{r}
eye.distance.mod <- aov(time~distance, data=effect)
summary.aov(eye.distance.mod)
```
Blocking factor decreases the MSE value, so it is effective

### c. Perform TukeyHSD on “Distance”. Which Distances bring the longest/ shortest focus time

```{r}
tukey.distance <- TukeyHSD(eye.distance.mod)
tukey.distance
plot(tukey.distance)
```


### d.  Calculate the sample size for power > .90, use power.anova.test().

```{r}
MSE <-1.275
trt.means <-tapply(time, list(subject, distance), mean)
power.anova.test(groups=5, n=2, between.var = var(trt.means), within.var = MSE, 
                 sig.level = 0.05)
```

```{r echo=FALSE}
detach()
```
***

## Question 3
### a. Setup a dataframe manually to use the Latin square design
```{r}

setup_q_3<- function () 
{
operator <- factor(rep(c(1,2,3,4), time=4), level=c(1,2,3,4), 
                   label=c("Op1", "Op2", "Op3", "Op4"))
order <- factor(rep(c(1,2,3,4), each=4), level=c(1,2,3,4), 
                label=c("Order1","Order2","Order3","Order4" ))
r1 <- c('C','D','A','B')
r2 <- c('B','C','D','A')
r3 <- c('A','B','C','D')
r4 <- c('D','A','B','C')
treat <- c(r1,r2,r3,r4)
time <- c(10,14,7,8,7,18,11,8,5,10,11,9,10,10,12,14)

experiment <-data.frame(operator, order, treat, time)
return (experiment)
}
experiment <- setup_q_3()
str(experiment)
head(experiment)
attach(experiment)
```

### b. Build a regression model, using aov(). Do the treatment affect the assembly time?

```{r}

time.aov <- aov(time~treat+order+operator)
summary(time.aov)
```
The pvalue for treatment is significant, so it does affect the assmebly time.

### c. Find the lowest assembly time.

```{r}
tukey <- TukeyHSD(time.aov)
tukey
plot(tukey)
```
A has smallest assembly time

```{r echo=FALSE}
detach()
```

***