---
title: "Lab2"
author: "Jas Sur"
date: "2/07/2022"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q.1
### a. Set up the data frame.

```{r p.1}

commercial_plastic_wrap <- c(7.66, 6.98, 7.80)
vacuum_packaged <- c(5.26, 5.44, 5.80)

#1% CO,40% O2, 59% N
co_o2_n<- c(7.41, 7.33, 7.04)
co2_100_percent <-c(3.51, 2.91, 3.6)

logcount <- c(commercial_plastic_wrap, vacuum_packaged, co_o2_n, co2_100_percent)

package <- rep(c("comm_pwrap", "vacuum_pkd", "co_o2_n", "co2"),
              each=3)
package <-factor(package)
bacteria  <-data.frame(cbind(logcount, package))
head(bacteria)
```


### b. Perform a stripchart, with line connecting means, of logcount vs package

```{r p.1.b}
stripchart(logcount~package, vertical=TRUE, pch=16, main="Stripchart")

logcount.means<-tapply(logcount, package, mean)
lines(logcount.means)

boxplot(logcount~package)
```

It looks like the co-o2-n and common_package_wrap has the same log count.


### c. Build a linear model, using aov() response as logcount. Do a summary.lm() 

```{r p.1.c}
model<- aov(logcount ~ package)
summary.aov(model)
summary.lm(model)
```
pvalue is smaller than 0.05 so it's evident that the means of some groups is not the same


### d. Perform a Bartlett test of equal variances.

```{r p.1.d}
logcount.aov<- aov(model)
res <- residuals(logcount.aov)
qqnorm(res, main="normality")
qqline(res)
bartlett.test(res~package)
```


The variance is same for different groups

### e. Perform a multiple comparison of treatment mean, using TukeyHSD()
```{r p.1.e}
tuskey <-TukeyHSD(model, "package")
tuskey
plot(tuskey)
```

## Q.2. Data: Tensile strength of Portland Cement

### a. Set up a data frame , with varibles: mixing (factor) and strength (response)

```{r p.2.a}
one <- c(3129, 3000, 2865, 2890)
two <- c(3200, 3300, 2975, 3150)
three <- c(2800, 2900, 2985, 3050)
four <- c(2600, 2700, 2600, 2765)
strength <- c(one, two, three, four)

material <- rep(c("1", "2", "3", "4"), each=4)
material <-factor(material)
cement  <-data.frame(cbind(strength, material))
head(cement)
```

### b. Perform a stripchart. Perform a Box plot.

```{r p.2.b}
stripchart(strength~material, vertical=TRUE, pch=16, main="Stripchart")

strength.means<-tapply(strength, material, mean)
lines(strength.means)

```

### c. Test the hypothesis that mixing techniques affect the strength of the cement. Use Î±=0.05
### What test do use. Perform the test. Conclusion.
H0 = mixing techniques doesn't affect the strength of material
Ha = mixing techniques affect the strength of material

assumptions
> data is normal

```{r p.2.c}
strength.aov.model<- aov(strength ~ material)
summary.aov(strength.aov.model)
summary.lm(strength.aov.model)


res <- residuals(strength.aov.model)
qqnorm(res, main="normality")
qqline(res)
shapiro.test(res)

#check variance
bartlett.test(res~material)
```

pvalue is greater than alpha (0.05), reject null hypothesis, accept alternative hypothesis that material does affect the strength 

### d. Use the Fisher LSD (Least Significant Difference) ð›¼ = 0.05 to make comparison
```{r p.2.d}
#install.packages("agricolae")
library(agricolae)

MSerror <- 12826
Fisher<- LSD.test(strength.aov.model, "material", MSerror, console=T)
Fisher

plot(Fisher)
```

Only one group has pvalue > 0.05 and rest of the groups has pvalue < 0.05,
so the means for each of other groups is not the same
comm_pwrap-co_o2_n     0.22 -0.6536808  1.0936808 0.8497833

## Q 3 

```{r p.3.data}
one <- c(143, 141, 150, 146)
two <- c(152, 149, 137, 143)
three <- c(134, 136, 132, 127)
four <- c(129, 127, 132, 129)

conductivity <- c(one, two, three, four)

coating <- rep(c("1", "2", "3", "4"), each=4)
coating <-factor(coating)
television  <- data.frame(cbind(conductivity, coating))
head(television)
```


### a. Is there a difference in conductivity due to coating type?
### alpha = 0.05

```{r p.3.a}
cond.model<- aov(conductivity ~ coating)
summary.aov(cond.model)
summary.lm(cond.model)

res <- residuals(cond.model)
qqnorm(res, main="normality")
qqline(res)
shapiro.test(res)
```

pvalue is greater than 0.05,so it's ascertained that the data is normal


compare the variance
> H0 = there is no difference in conductivity var1 = var2
> Ha = there is a difference in conductivity var1 <> var2

```{r p.3.a.1}
bartlett.test(res~coating)
```


pvalue is greater than 0.05, so null hypothesis is accepted and it's ascertained  the variance is same

Since the data is normal and variance is equal, it may be concluded that 
there is no difference in conductivity due to coating type


### b. Estimate the mean and the treatment effects view the model output

```{r p.3.b}
stripchart(conductivity~coating, vertical=TRUE, pch=16, main="Stripchart")

cond.means<-tapply(conductivity, coating, mean)
lines(cond.means)
```

It can be seen from the stripchart that coating type of 1,2 have the same conductivity while 3, 4 have another group of same conductivity, but doesn't match 1,2

The pvalue from aov test is smaller than 0.05, so we have sufficient evidence to say that one of the means is different from the others.
```{r p.3.c}
tukey.95 <- TukeyHSD(cond.model, "coating")

tukey.95

plot(tukey.95)
```

For coating type 4, the confidence intervals for the mean value 
between groups 4-2 and 4-1 contain the value zero, 
which indicates that there is a statistically significant difference 
in mean loss between the two groups. 
This is consistent with the fact that two of these groups of the 
p-values from our hypothesis tests are below 0.05.

```{r p.3.c.1}
tukey.99 <- TukeyHSD(cond.model, "coating", conf.level = 0.99)
tukey.99

plot(tukey.99)
```

For coating type 4, the confidence intervals for the mean value 
between groups 4-2 and 4-1 contain the value zero, 
which indicates that there is a statistically significant difference 
in mean loss between the two groups. 
This is consistent with the fact that two of these groups of the 
p-values from our hypothesis tests are below 0.05.

There is not much statistical difference between .95 and .99 

### d. Test all pairs of mean using Fisher LSD
```{r p.4.d}
MSerror <-19.69 
Fisher<- LSD.test(cond.model, "coating", MSerror, console=T)
Fisher

plot(Fisher)
```

2,1 (group a) and 3,4 (group b) have same mean

### e. Assuming that coating type 4 is currently in use, what are your recommendations to the manufacturer? We wish to minimize conductivity.
Since group 3 and 4 have the same conductivity, coating# 4 can be used interchangebly with coating# 3.