---
title: "midterm-new"
author: "Jas Sur"
date: "3/21/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 9.

```{r}
Score<-c(23,24,25,36,35,36,28,24,27,27,28,26,34,38,39,35,35,34,31,32,29,33,34,
         35,26,27,25,24,23,28,37,39,35,26,29,25,38,36,35,34,38, 
         36,36,37,34,34,36,39,34,36,31,28,26,24)

Temperature.num<-c(300,300,300,300,300,300,300,300,300,300,300,300,
               300,300,300,300,300,300,300,300,300,300,300,300,
               300,300,300,350,350,350,350,350,350,350,350,350,
               350,350,350,350,350,350,350,350,350,350,350,350,
               350,350,350,350,350,350)

Operator.num<-c(1,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,2,3,3,3,3,3,3,3,3,
                3,1,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,2,3,3,3,3,3, 3,3,3,3) 

CycleTime.num<-c(40,40,40,50,50,50,60,60,60,40,40,40,50,50,50,60,60,60,40,40,40,
                 50,50,50,60,60,60,40,40,40,50,50,50,60,60,60,40,40,40,50,50,50,
                 60,60,60,40,40,40,50,50,50,60,60,60)
fabrics<-data.frame(Score,Temperature.num,Operator.num)
fabrics$Temperature<-as.factor(Temperature.num)
fabrics$Operator<-as.factor(Operator.num)
fabrics$CycleTime<-as.factor(CycleTime.num)

attach(fabrics)
par(mar=c(1, 1, 1, 1))
#par(mfrow=c(2,2))
interaction.plot(Temperature,Operator,Score)
interaction.plot(CycleTime,Operator,Score)
interaction.plot(CycleTime,Temperature,Score)
detach()
```
Operator and Temperature has no interaction
CycleTime and Operator has some interaction
CycleTime and Temperature has some interaction

### a. Build a regression model. Any interaction among factors?

```{r}
fabrics.model<-aov(Score~Temperature*CycleTime*Operator )
summary.aov(fabrics.model)
```
Temperature:Operator 0.193895  is not significant



### b. Which factor can be used as Blocking factor? Is the Blocking factor effective?

```{r}
fabrics.model.nonblock<-aov(Score~Temperature+CycleTime)
summary.aov(fabrics.model.nonblock)
```
MSE is increased, so Operator is a valid block.

### c. Consider any reduced model?
```{r}
fabrics.model<-aov(Score~Temperature+CycleTime*Operator )
summary.aov(fabrics.model)
```
All factors are significant.
